### 2017-06-16 paper_revised/referee_reply.txt
-Work on the reply.

Dear A&A Editor,

Thank you for sending us the report of the referee, whom we thank for
carefully reading our manuscript and providing very useful feedback. We 
have made several changes addressing the issues raised in the report, 
which are listed below together with our response. The changes made in 
the manuscript are highlighted in boldface, to facilitate identification. 
We hope that with these amendments the paper will be considered 
acceptable for publication on Astronomy and Astrophysics.

       Best wishes,

       The authors

----------------------------------------------------------------------
Report of the Referee -- AA/2017/30935
----------------------------------------------------------------------

> The paper presents a new procedure to construct sequences of equilibrium
> tori around Kerr black holes that are endowed with a toroidal magnetic
> field. These configurations are not expected to be produced under generic
> conditions but the construction of such equilibria is interesting per se
> and has a rather long history. Furthermore, as discussed by the authors,
> the availability of such equilibria can be employed in test simulations
> of general-relativistic codes in ideal MHD.

> The paper is nicely written with a clear discussion of the procedure
> followed and of the properties of the tori that have been found. The
> paper is essentially ready for publication but there are a few aspects
> that could be improved and that I ask the authors to consider.
>

> - Title: I frankly find it unnecessarily colloquial. I do not insist but
> I think a more pragmatic title such as "A new method for the
> construction of magnetized tori around black holes" is scientifically
> more informative.

We agree with the referee about the title being colloquial. Our choice was
motivated by the title of the paper of Qian et al (2009), “The Polish 
doughnuts revisited …”, of which our paper might be considered an extension 
(by incorporating magnetic fields). We would prefer to keep the original 
title, if the referee concurred. If that were not the case, we would not
object to having it changed to his/her suggestion. (“A new method for the
construction of magnetized tori around black holes”.)

> - Sec. 2.1: the authors introduce the concept of cusp but do not really
> define it

> - Sec. 2.1: they introduce the equation of state (13) but do not discuss
> how they treat the specific internal energy that is part of the
> enthalpy used in (13). A comment should be added.
>

> - Sec. 2.1: I am not clear how the magnetic field is actually
> prescribed. The author mention that the only nonzero component of
> the magnetic field in the comoving frame is the azimuthal one but I am
> still unclear how the strength of this component is specified. A
> comment should be added.

> - Sec. 2.1: More on this topic: do the fluid isobaric surfaces (i.e. the
> surfaces of constant fluid pressure) coincide with the magnetic
> isobaric surfaces (i.e. the surfaces of constant magnetic pressure)?
> How are these surfaces related to the iso-density surfaces? A comment
> should be added.
>

> - Sec. 3: the authors are a bit too sloppy when referring to the fluid
> pressure, which they call respectively: "fluid pressure", "gas
> pressure" "thermal pressure", and sometimes simply as "pressure". Since
> distinguishing the fluid, magnetic and total pressure is important, I
> ask the authors to always just talk of "fluid pressure".

We have followed the recomendation of the referee and in the revised version
of the paper only mentions to “fluid pressure”, “magnetic pressure”, and 
“total pressure” appear. Those appearances are highlighted in boldface
throughout.

> - Sec. 3: similar considerations apply to the density, which should
> be referred to as "rest-mass density"
>

Done.

> - Sec. 3: the authors discuss that the rest-mass density distribution can
> "be trivially obtained inverting the barotropic relation, p = K
> rho^kappa". I am not sure this is so trivial unless an assumption is
> made on the specific internal energy; the authors should discuss more
> clearly how they actually obtain the density from Eqs. (12) and (14).
>

> - Sec. 3: the authors discuss that Eq. (24) diverges for theta --> pi/2,
> i.e. at the equator. I am a bit puzzled by this statement and I wonder if
> this is an exact statement, ie whether it is true that

> partial_theta W = 0 for theta --> pi/2.

> If this is an exact result, I urge the authors to prove it. If it is
> not, I urge the authors to simply mention that the gradients become
> very small.
>

> - Fig. 1: as with the middle panel, I suggest that the right
> panel should also show the logarithm of the rest-mass density.
>

> - Fig. 2: all of the panels would benefit from a legend showing the
> corresponding values of gamma and beta. This would avoid having to read the caption repeatdly. The same is true also for Fig. 3.
>

> - Fig. 4: the caption should be corrected in "beta = 10^3 --> beta_{m_c}
> = 10^3".

> - Fig. 4: the authors do not discuss what is the largest magnetization
> that they can handle. I agree that beta_{m_c} = 10^{-3} is probably
> already quite high, but what happens for say beta_{m_c} = 10^{-6}? Can
> a solution still be found?

> - Sec. 4: I find the discussion in this section the less clear; a lot
> of the considerations appear phenomenological but I can't find the
> simple back-of-the-envelope demonstration that what the authors observe
> numerically is what should be expected on physical grounds. For
> example, the authors find that the rho_max increases with magnetization
> (left panel of Fig. 6) or that the position of the rest-mass density
> maximum moves in with increasing magnetization (middle panel of
> Fig. 6). What are the physical reasons for this behavior? I have some
> suggestions on this but I urge the authors to provide some simple and
> intuitive explanations first.
>

> - References: the list is already quite complete but maybe it can be
> complemented with the following ones, many of which are actually
> written by one of the authors.

> A comprehensive discussion of the properties of tori from merging binaries:

> http://adsabs.harvard.edu/abs/2016arXiv160703540B
>

> First detailed discussion of non-constant specific angular momentum tori:

> http://adsabs.harvard.edu/abs/2004MNRAS.354.1040M

> First discussion of the dynamics of magnetized tori built with the
> Komissarov solution

> http://adsabs.harvard.edu/abs/2007MNRAS.378.1101M

We have incorporated in the article the first and the third references 
suggested by the referee. However, we have not incorporated the second 
reference. While this is a very relevant work in the context of oscillations 
of non-constant-angular momentum disks, including this reference in our 
work seems to us a bit forced since it discusses aspects which are a bit 
away from the main discourse of our paper.