### 2017-06-20 paper_revised/reply_referee.tex
added reply_referee.tex

%
% version 27.10.16
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%        reply
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{article}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{amsfonts}
\textheight 22.cm
\textwidth 16.5 cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\topmargin=-1.cm
\hoffset -0.5cm
 \usepackage{amssymb}

\tolerance=10000
\pagenumbering{arabic}
\textheight 22.cm
\textwidth 16.5 cm
\oddsidemargin 0.5cm\evensidemargin 0.5cm
\topmargin=-1.cm
\hoffset -0.5cm
\date{\today}

\newcommand{\bnabla}{\mbox{\boldmath $\nabla$}}
%\newcommand{\bomega}{\mbox{\boldmath $\omega$}}
\newcommand{\bOmega}{\mbox{\boldmath $\Omega$}}

\newcommand{\insertplot}[5]{\begin{figure}
 \hfill\hbox to 0.05in{\vbox to #5in{\vfill
 \inputplot{#1}{#4}{#5}}\hfill}
 \hfill\vspace{-.1in}
 \caption{#2}\label{#3}
 \end{figure}}
 \newcommand{\inputplot}[3]{\special{ps:: gsave #2 -36 mul 0 rmoveto
   currentpoint translate #2 7.0 div #3 5.0 div scale}
 \special{ps: plotfile #1}\special{ps:: grestore}}
\newcounter{fig}   \newcommand{\lbfig}[1]{\refstepcounter{fig}
\label{#1} }

\newcommand{\vphi}{\varphi}
\newcommand{\DS}{\displaystyle}
\newcommand{\pih}{\frac{\pi}{2}}
\newcommand{\sqdetg}{\sqrt{-g}}
\newcommand{\sqdetgi}{\frac{1}{\sqrt{-g}}}
\newcommand{\edil}{e^{2\kappa \phi}}
\newcommand{\bA}{\bar{A}}
\newcommand{\bF}{\bar{F}}
\newcommand{\bD}{\bar{D}}

\renewcommand{\a}{\alpha}
\renewcommand{\b}{\beta}
\renewcommand{\c}{\gamma}
\renewcommand{\d}{\delta}  
\newcommand{\g}{\nu}
\renewcommand{\l}{\lambda}
\renewcommand{\t}{\theta}
\newcommand{\rot}{\mbox{rot}}
\renewcommand{\div}{\mbox{div}}
\newcommand{\cosec}{\mbox{cosec}}

\newcommand{\ta}{\theta}
\newcommand{\Si}{\Sigma}
\newcommand{\vf}{\varphi}
\newcommand{\dd}{\mbox{d}}
\newcommand{\tr}{\mbox{tr}}
\newcommand{\la}{\lambda}
\newcommand{\ka}{\kappa}
\newcommand{\f}{\phi}
\newcommand{\al}{\alpha}
\newcommand{\ga}{\gamma}
\newcommand{\de}{\delta}
\newcommand{\si}{\sigma}
\newcommand{\bomega}{\mbox{\boldmath $\omega$}}
\newcommand{\bsi}{\mbox{\boldmath $\sigma$}}
\newcommand{\bchi}{\mbox{\boldmath $\chi$}}
\newcommand{\bal}{\mbox{\boldmath $\alpha$}}
\newcommand{\bpsi}{\mbox{\boldmath $\psi$}}
\newcommand{\brho}{\mbox{\boldmath $\varrho$}}
\newcommand{\beps}{\mbox{\boldmath $\varepsilon$}}
\newcommand{\bxi}{\mbox{\boldmath $\xi$}}
\newcommand{\bbeta}{\mbox{\boldmath $\beta$}}
\newcommand{\ee}{\end{equation}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\be}{\begin{equation}}
\newcommand{\bea}{\begin{eqnarray}}

\newcommand{\pont}{{\,^\ast\!}R\,R}

\newcommand{\ii}{\mbox{i}}
\newcommand{\e}{\mbox{e}}
\newcommand{\pa}{\partial}
\newcommand{\Om}{\Omega}
\newcommand{\vep}{\varepsilon}
%\newcommand{\bfR}{{\bf R}}
\newcommand{\bfph}{{\bf \phi}}
\newcommand{\lm}{\lambda}
\def\theequation{\arabic{equation}}
%\topmargin= -02cm\textheight= 23.cm\textwidth= 16.cm
%\oddsidemargin=-01cm\evensidemargin=-01cm \font\sqi=cmssq8
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\newcommand{\re}[1]{(\ref{#1})}
\newcommand{\R}{{\rm I \hspace{-0.52ex} R}}
\newcommand{\N}{{\sf N\hspace*{-1.0ex}\rule{0.15ex}%
{1.3ex}\hspace*{1.0ex}}}
\newcommand{\Q}{{\sf Q\hspace*{-1.1ex}\rule{0.15ex}%
{1.5ex}\hspace*{1.1ex}}}
\newcommand{\C}{{\sf C\hspace*{-0.9ex}\rule{0.15ex}%
{1.3ex}\hspace*{0.9ex}}}
\newcommand{\eins}{1\hspace{-0.56ex}{\rm I}}
\renewcommand{\thefootnote}{\arabic{footnote}}

\def\theequation{\thesection.\arabic{equation}}

\begin{document}

\title{ 
{\bf Reply to the  Referee report:} \\ {\large AA/2017/30935} \\  {\large  \textit{Magnetised Polish doughnuts revisited}\\ by Sergio Gimeno-Soler and Jos\'e A. Font.}
}

\maketitle

Dear A\&A Editor,

\bigskip

Thank you for sending us the report of the referee, whom we thank for
carefully reading our manuscript and providing very useful feedback. We 
have made several changes addressing the issues raised in the report, 
which are listed below together with our response. The changes made in 
the manuscript are highlighted in boldface, to facilitate identification. 
We hope that with these amendments the paper will be considered 
acceptable for publication on \textit{Astronomy and Astrophysics}.

\bigskip

Best wishes,

\bigskip

The authors

\bigskip

\section*{Referee report -- AA/2017/30935}

We are grateful to the referee for the interesting comments provided in his/her report.
We have followed the referee's suggestions and have incorporated most of his/her comments 
in our manuscript. Below, we ({\bf A}) address the various issues raised by the referee ({\bf R}).

\bigskip

{\bf R:} The paper presents a new procedure to construct sequences of equilibrium
tori around Kerr black holes that are endowed with a toroidal magnetic
field. These configurations are not expected to be produced under generic
conditions but the construction of such equilibria is interesting per se
and has a rather long history. Furthermore, as discussed by the authors,
the availability of such equilibria can be employed in test simulations
of general-relativistic codes in ideal MHD.The paper is nicely written with a clear discussion of the procedure
followed and of the properties of the tori that have been found. The
paper is essentially ready for publication but there are a few aspects
that could be improved and that I ask the authors to consider.

\bigskip

{\bf R:} - Title: I frankly find it unnecessarily colloquial. I do not insist but
I think a more pragmatic title such as ``A new method for the
construction of magnetized tori around black holes" is scientifically
more informative.

\bigskip

{\bf A:} We agree with the referee about the title being colloquial. Our choice was
motivated by the title of the paper of Qian et al (2009), ``The Polish 
doughnuts revisited …", of which our paper might be considered an extension 
(by incorporating magnetic fields). We would prefer to keep the original 
title, if the referee concurred. If that were not the case, we would not
object to having it changed to his/her suggestion. (``A new method for the
construction of magnetized tori around black holes".)

\bigskip

{\bf R:} - Sec. 2.1: the authors introduce the concept of cusp but do not really
define it

\bigskip

{\bf A:} We have included the following definition: ``The cusp is defined as the circle 
in the equatorial plane on which the pressure gradient vanishes and the angular 
momentum of the disk equals the Keplerian angular momentum."

\bigskip

{\bf R:} - Sec. 2.1: they introduce the equation of state (13) but do not discuss
how they treat the specific internal energy that is part of the
enthalpy used in (13). A comment should be added.

\bigskip

{\bf A:}

\bigskip

{\bf R:} - Sec. 2.1: I am not clear how the magnetic field is actually
prescribed. The author mention that the only nonzero component of
the magnetic field in the comoving frame is the azimuthal one but I am
still unclear how the strength of this component is specified. A
comment should be added.

\bigskip

{\bf A:}

\bigskip

{\bf R:} - Sec. 2.1: More on this topic: do the fluid isobaric surfaces (i.e. the
surfaces of constant fluid pressure) coincide with the magnetic
isobaric surfaces (i.e. the surfaces of constant magnetic pressure)?
How are these surfaces related to the iso-density surfaces? A comment
should be added.

\bigskip

{\bf A:}

\bigskip

{\bf R:} - Sec. 3: the authors are a bit too sloppy when referring to the fluid
pressure, which they call respectively: ``fluid pressure", ``gas
pressure" ``thermal pressure", and sometimes simply as ``pressure". Since
distinguishing the fluid, magnetic and total pressure is important, I
ask the authors to always just talk of ``fluid pressure".

\bigskip

{\bf A:} We have followed the recomendation of the referee and in the revised version
of the paper only mentions to ``fluid pressure", ``magnetic pressure", and 
``total pressure" appear. Those appearances are highlighted in boldface
throughout.

\bigskip

{\bf R:} - Sec. 3: similar considerations apply to the density, which should
be referred to as ``rest-mass density"

\bigskip

{\bf A:} Done.

\bigskip

{\bf R:} - Sec. 3: the authors discuss that the rest-mass density distribution can
``be trivially obtained inverting the barotropic relation, $p = K
rho^kappa$". I am not sure this is so trivial unless an assumption is
made on the specific internal energy; the authors should discuss more
clearly how they actually obtain the density from Eqs. (12) and (14).

\bigskip

{\bf A:}

\bigskip

{\bf R:} - Sec. 3: the authors discuss that Eq. (24) diverges for $\theta \rightarrow \pi/2$,
i.e. at the equator. I am a bit puzzled by this statement and I wonder if
this is an exact statement, ie whether it is true that

\begin{equation}
\partial_{\theta} W = 0 \ \mathrm{for} \ \theta \rightarrow \frac{\pi}{2}.
\end{equation}

If this is an exact result, I urge the authors to prove it. If it is
not, I urge the authors to simply mention that the gradients become
very small.

\bigskip

{\bf A:}

\bigskip

{\bf R:} - Fig. 1: as with the middle panel, I suggest that the right
panel should also show the logarithm of the rest-mass density.

\bigskip

{\bf A:}

\bigskip

{\bf R:} - Fig. 2: all of the panels would benefit from a legend showing the
corresponding values of gamma and beta. This would avoid having to read the caption repeatdly. The same is true also for Fig. 3.

\bigskip

{\bf A:}

\bigskip

{\bf R:} - Fig. 4: the caption should be corrected in ``$\beta = 10^3$ --> $\beta_{m_c}
= 10^3$".

\bigskip

{\bf A:}

\bigskip

{\bf R:} - Fig. 4: the authors do not discuss what is the largest magnetization
that they can handle. I agree that $\beta_{m_c} = 10^{-3}$ is probably
already quite high, but what happens for say $\beta_{m_c} = 10^{-6}$? Can
a solution still be found?

\bigskip

{\bf A:}

\bigskip

{\bf R:} - Sec. 4: I find the discussion in this section the less clear; a lot
of the considerations appear phenomenological but I can't find the
simple back-of-the-envelope demonstration that what the authors observe
numerically is what should be expected on physical grounds. For
example, the authors find that the $\rho_{\mathrm{max}}$ increases with magnetization
(left panel of Fig. 6) or that the position of the rest-mass density
maximum moves in with increasing magnetization (middle panel of
Fig. 6). What are the physical reasons for this behavior? I have some
suggestions on this but I urge the authors to provide some simple and
intuitive explanations first.

\bigskip

{\bf A:}

\bigskip

{\bf R:} - References: the list is already quite complete but maybe it can be
complemented with the following ones, many of which are actually
written by one of the authors.

A comprehensive discussion of the properties of tori from merging binaries:

http://adsabs.harvard.edu/abs/2016arXiv160703540B

First detailed discussion of non-constant specific angular momentum tori:

http://adsabs.harvard.edu/abs/2004MNRAS.354.1040M

First discussion of the dynamics of magnetized tori built with the Komissarov solution

http://adsabs.harvard.edu/abs/2007MNRAS.378.1101M

\bigskip

{\bf A:} We have incorporated in the article the first and the third references 
suggested by the referee. However, we have not incorporated the second 
reference. While this is a very relevant work in the context of oscillations 
of non-constant-angular momentum disks, including this reference in our 
work seems to us a bit forced since it discusses aspects which are a bit 
away from the main discourse of our paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
\begin{small}
\begin{thebibliography}{99}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
%

%  
\end{thebibliography}

 \end{small}

 \end{document}