### 2017-11-14 magnetised_polish_doughnuts_revisited/paper_revised_v2/referee_reply.txt
new paper added

Referee Report

The authors have followed carefully all of the suggestions made and I am
pleased to see that they were able to provide more rigorous support to
some of the arguments presented in the paper. The new appendices A and B
present valuable additions to the paper.

In view of these considerations I am glad to recommend the publication of
the manuscript as a regular paper on A&A. At the same time, I'd like the
authors to consider some of the small suggestions listed below and
addressing some of the questions raised.

## A. ...We have included the following comment at the end of the paragraph
## following Eq. (27): "We note that the fluid enthalpy w includes the
## rest-mass density rhos, and can thus be defined as w = rho h where h
## is the specific enthalpy. The relativistic definition of this quantity
## is h = 1+epsilon+p/rho, where epsilon is the specific internal
## energy. From a thermodynamical point of view a non-relativistic fluid
## satisfies h = 1. Therefore, since we use poytropic equations of state
## (relating p with either w or rho in the same functional form) we are
## implictly assuming that h = 1, i.e. the discs we build in our
## procedure are non-relativistic from a thermodynamical point of view."

The authors should note that "w" is the *enthalpy density* and not the
enthalpy. This becomes clear: any "specific" quantity becomes a
"density" when multiplied by a rest-mass density.

## A: The reason we show the middle panel in logarithm scale and the
## right panel in linear scale is because this is the choice in
## Komissarov (2006). We have followed the same choice in order to
## facilitate the qualitative and quantitative comparison for this
## constant angular momentum test. If the referee concurred, we would
## prefer to keep the right panel in linear scale. If not, we would not
## mind changing the right panel to logarithmic scale.

Following too closely the work of Komissarov is not necessary and not
necessarily a good idea. Rather, I think that a plot in logarithmic
scale would be more informative.

## A: We have built models with very small values of beta_mc, such as
## beta_mc = 1e-20 (extremely high magnetisation) without encountering
## numerical difficulties. In practice we have found that there are no
## qualitative differences in the structure of the discs once beta_mc
## becomes smaller than a sufficiently small value...

It is important to incorporate this comment in the revised version.

## R: - Fig. 2: all of the panels would benefit from a legend showing the
## corresponding values of gamma and beta. This would avoid having to read
## the caption repeatedly. The same is true also for Fig. 3.
##
## A: Done.

I think that the legend should also report the value of the
black-hole's spin as this makes the various panels readable also
without having to consult the caption.

## If the referee considered these explanations satisfactory and would
## like to have them included in the manuscript, we would be happy to do
## so.

I find the explanation satisfactory and, as mentioned above, I find
Appendix B a nice addition to the paper.