### 2017-11-14 magnetised_polish_doughnuts_revisited/paper_revised/beta_gamma_output.tex

\documentclass{aa}

\usepackage{graphicx}
\usepackage{txfonts}
\usepackage{hyperref}
\usepackage{color}
\usepackage{float}

\begin{document}

\title{beta and gamma output}

\author{Sergio Gimeno-Soler\inst{1}}
\institute{Departamento de Astronom\'{\i}a y Astrof\'{\i}sica, Universitat de Val\`encia, Dr. Moliner 50, 46100, Burjassot (Val\`encia), Spain.\\\email{sergio.gimeno@uv.es}
}

   \date{}

  \abstract
{Document to capture value of $\beta$ and $\gamma$ for the graphs}

   \keywords{accretion, accretion discs -- black hole physics -- magnetohydrodynamics (MHD)}

   \maketitle

%%%%%%%%%%%
\section{Introduction}

\begin{equation}
\beta = 0
\end{equation}
\begin{equation}
\gamma = 0
\end{equation}
\end{document}